### 2021-06-16 sec_summary.tex

\section{Summary}
\label{sec:summary}

In summary, a quantitative study of long-range two-particle correlations in \pythia events with the string shoving model has been performed.
Clear long-range near-side correlations are observed in the string shoving model, whereas no such structure is seen in \pythia events with the default configuration.
In the comparison with experimental data, the string shoving model qualitatively describes long-range correlations in high multiplicity events.
However, one difference is that long-range near-side correlations exist even at low multiplicity region in the string shoving model.
Due to this fact, the model overestimate associated yields of long-range near-side correlation functions, and flow coefficients ($v_{2,2}$) obtained from the template fit method are lower than
the experimental results.
%expected values with low multiplicity templates of no long-range near-side correlations.
A method to subtract jet correlations in associated yields is tested, and the subtracted yields of charged hadrons for $1<\pt<2~\mathrm{GeV}/c$ shows a similar multiplicity dependence with the experimental results.